TGI-spinal-power

### 2022-05-30 analysis/tgispine_zoib_power-sim.Rmd
updating power calc simulation script, changing to beta distrib

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(brms)
library(modelsummary)
library(tibble)
library(DHARMa)
library(glmmTMB)
library(reshape2)
library(truncnorm)
library(tidyverse)
library(plyr)
library(data.table)

@@

# transform variables into proportions (aka divide by 100), this makes the effect size estimates more logical
df_VAS$beta <- df_VAS$beta/100
df_VAS$ID <- factor(df_VAS$ID)

# Summarise beta per rating, per condition, per participant
# median
df_sum_pp <- aggregate(beta~ID*manipulation*condition*cold_location*VAS, 
                    median, data = df_VAS)

@@

# creating individual data-frames for specific conditions (to make simulation easier)
cnt_cold <- df_sum[df_sum$manipulation == 'CNT' & df_sum$VAS == 'cold' ,]
cnt_warm <- df_sum[df_sum$manipulation == 'CNT' & df_sum$VAS == 'warm' ,]
cnt_burn <- df_sum[df_sum$manipulation == 'CNT' & df_sum$VAS == 'burn' ,]
tgi_cold <- df_sum[df_sum$manipulation == 'TGI' & df_sum$VAS == 'cold' ,]
tgi_warm <- df_sum[df_sum$manipulation == 'TGI' & df_sum$VAS == 'warm' ,]
tgi_burn <- df_sum[df_sum$manipulation == 'TGI' & df_sum$VAS == 'burn' ,]
```

# Simulating data function
```{r}
simulate_subj <- function(subid, llim, ulim) {
# trial info
trialN = 12
conditionN = 8
ratingN = 3
# creating simulation data-frame
df <- data.frame(matrix(ncol = 0, nrow = trialN*conditionN*ratingN))
df$id <- rep(subid,trialN*conditionN*ratingN)
df$trialn <- rep(1:trialN,conditionN*ratingN)
df$dermatome <- rep(rep(c(rep("within",trialN*2), rep("across",trialN*2)),2),ratingN)
df$cold_location <- rep(rep(c(rep("dist_rostr",trialN), rep("prox_caud", trialN)),conditionN/2),ratingN)
df$stimuli <- rep(c(rep("tgi",trialN*conditionN/2), rep("non-tgi",trialN*conditionN/2)),ratingN)
df$quality <- c(rep("cold",trialN*conditionN), rep("warm",trialN*conditionN), rep("burn",trialN*conditionN))

@@

power_c
```

# Then warm
```{r}
# the warm model has the smallest effect size, so should extract power calculation for this model only
# do this for 40 participants, for 1000 simulations
nSims <- 1000
n <- 32 # number of participants to simulate
po <- data.frame()
for (i in 1:nSims){
  # creating individual data-frames from simulated data
  df_sim <- data.frame()
  for (i in 1:n) {
    df <- simulate_subj(i, 0, 1)
    df_sim <- rbind(df_sim,df)
    }
  # then run the model based on simulated data
  m <- model(df_sim, 'warm')
  m1 <- summary(m)
  # extracting the p-value and effect size
  tmp <- c(m1$coefficients$cond[4,4], m1$coefficients$cond[4,1])
  po <- rbind(po,tmp)
}

# rename columns
names(po)[1] <- 'pval'
names(po)[2] <- 'eff'
po$sig <- as.numeric(po$pval < .05) #is significant
po$sig[is.na(po$sig)] <- 0 #removing any NAs (code as unsig)

power <- sum(po$sig)/nSims
# some renaming
po_w <- po
power_w <- power

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(brms)
library(modelsummary)
library(tibble)
library(DHARMa)
library(glmmTMB)
library(reshape2)
library(truncnorm)
library(tidyverse)
library(plyr)
library(data.table)
library(extraDistr)

@@

# transform variables into proportions (aka divide by 100), this makes the effect size estimates more logical
df_VAS$beta <- df_VAS$beta/100
df_VAS$ID <- factor(df_VAS$ID)
```

# Summary statistics
```{r}
# Summarise beta per rating, per condition, per participant
# median
df_sum_pp <- aggregate(beta~ID*manipulation*condition*cold_location*VAS, 
                    median, data = df_VAS)

@@

# creating relevant data-frames for simulation
tgi_cold <- df_sum %>% 
  filter(manipulation == 'TGI', VAS == 'cold')
tgi_warm <- df_sum %>% 
  filter(manipulation == 'TGI', VAS == 'warm')
tgi_burn <- df_sum %>% 
  filter(manipulation == 'TGI', VAS == 'burn')
cnt_cold <- df_sum %>% 
  filter(manipulation == 'CNT', VAS == 'cold')
cnt_warm <- df_sum %>% 
  filter(manipulation == 'CNT', VAS == 'warm')
cnt_burn <- df_sum %>% 
  filter(manipulation == 'CNT', VAS == 'burn')

# finally, count how many zeros there are in the data
# this is to model zero inflated regression
# cold cnt
CNTc_zero <- df_VAS %>% 
    filter(VAS == 'cold' & manipulation == 'CNT') %>% 
    dplyr::count(beta == 0) %>% 
    mutate(prop = n / sum(n))
# cold tgi
TGIc_zero <- df_VAS %>% 
    filter(VAS == 'cold', manipulation == 'TGI') %>% 
    dplyr::count(beta == 0) %>% 
    mutate(prop = n / sum(n))
# warm cnt
CNTw_zero <- df_VAS %>% 
    filter(VAS == 'warm', manipulation == 'CNT') %>% 
    dplyr::count(beta == 0) %>% 
    mutate(prop = n / sum(n))
# warm tgi
TGIw_zero <- df_VAS %>% 
    filter(VAS == 'warm', manipulation == 'TGI') %>% 
    dplyr::count(beta == 0) %>% 
    mutate(prop = n / sum(n))

# First, run a model on the pilot data, one for each VAS, to use for simulation
```{r}
# COLD
mc.og = glmmTMB::glmmTMB(
    beta ~ manipulation + condition + cold_location + trial_n + (1|ID),
    family = glmmTMB::beta_family(),
    ziformula = ~1+manipulation,
    data = filter(df_VAS, VAS == 'cold'),
    na.action = na.omit) 
# WARM
mw.og = glmmTMB::glmmTMB(
    beta ~ manipulation + condition + cold_location + trial_n + (1|ID),
    family = glmmTMB::beta_family(),
    ziformula = ~1+manipulation,
    data = filter(df_VAS, VAS == 'warm'),
    na.action = na.omit) 
# BURN
mb.og = glmmTMB::glmmTMB(
    beta ~ manipulation + condition + cold_location + trial_n + (1|ID),
    family = glmmTMB::beta_family(),
    ziformula = ~1+manipulation,
    data = filter(df_VAS, VAS == 'burn'),
    na.action = na.omit)

summary(mc.og)
summary(mw.og)
summary(mb.og)
```

# Data to use in simulation
```{r}

# Simulating data function - simulate model estimates (mean & sd) into beta distribution
```{r}
simulate_subj <- function(subid, llim, ulim) {
# trial info
trialN = 12
conditionN = 8
ratingN = 3
# creating simulation data-frame
df <- data.frame(matrix(ncol = 0, nrow = trialN*conditionN*ratingN))
df$id <- rep(subid,trialN*conditionN*ratingN)
df$trialn <- rep(1:trialN,conditionN*ratingN)
df$dermatome <- rep(rep(c(rep("within",trialN*2), rep("across",trialN*2)),2),ratingN)
df$cold_location <- rep(rep(c(rep("dist_rostr",trialN), rep("prox_caud", trialN)),conditionN/2),ratingN)
df$stimuli <- rep(c(rep("tgi",trialN*conditionN/2), rep("non-tgi",trialN*conditionN/2)),ratingN)
df$quality <- c(rep("cold",trialN*conditionN), rep("warm",trialN*conditionN), rep("burn",trialN*conditionN))

@@

power_c
write.csv(po_c, 'simulation_model_cold.csv', row.names = FALSE)
```

# Then warm
```{r}
# the warm model has the smallest effect size, so should extract power calculation for this model only
# do this for 40 participants, for 1000 simulations
nSims <- 1000
n <- 40 # number of participants to simulate
po <- data.frame()
for (i in 1:nSims){
  # creating individual data-frames from simulated data
  df_sim <- data.frame()
  for (i in 1:n) {
    df <- simulate_subj(i, 0, 1)
    df_sim <- rbind(df_sim,df)
    }
  # then run the model based on simulated data
  m <- model(df_sim, 'warm')
  m1 <- summary(m)
  # extracting the p-value and effect size
  tmp <- c(m1$coefficients$cond[4,4], m1$coefficients$cond[4,1])
  po <- rbind(po,tmp)
}

power <- sum(po$sig)/nSims
# some renaming
po_w <- po
write.csv(po_w, 'simulation_model_warm.csv', row.names = FALSE)
power_w <- power
